---
title: "NRL analysis"
output: html_document
---

```{r setup, include=FALSE}
library(broom)
library(rio)
library(RCurl)
library(readxl)
library(httr)
library(tidyverse)
library(plotly)
library(corrplot)
library(ggcorrplot)
library(stringr)
library(vip)
library(caTools)
library(rstanarm)
library(tidymodels)
library(DT)
# library(devtools)

# install("C:/Users/dan.fraser/Documents/NRLfastR", force = TRUE)
# library(NRLfastR)

# renv::restore()
```

All the models and insights are based off match data from foxsports.com.au for all the matches in four NRL seasons from 2021 to 2024, with a total of 828 games (201 in both 2021 and 2022, and 213 in both 2023 and 2024).

```{r data, include=FALSE}
# Importing data
NRL_seasons_Original <- read_xlsx("C:/Users/dan.fraser/Downloads/nrl_all_match_stats_2021_2024_wide.xlsx")

NRL_Endof2024 <- read_xlsx("C:/Users/dan.fraser/Downloads/nrl_all_match_stats_2024_wide.xlsx")

NRL_specificurls <- read_xlsx("C:/Users/dan.fraser/Downloads/nrl_match_stats_specific_urls.xlsx") %>% 
  mutate(ID = c('2023-25-4','2023-25-4','2024-8-2','2024-8-2'))

NRL_2021_2024 <- rbind(NRL_seasons_Original,NRL_Endof2024,NRL_specificurls)
```

```{r NRL_seasons, include=FALSE}
# Merging data/Creating NRL_seasons
NRL_seasons <- NRL_2021_2024 %>% rename("Set Restarts Awarded" = Awarded) %>%
  mutate(Possession = as.numeric(gsub("%", "",Possession)),
                                      Territory = as.numeric(gsub("%", "",Territory))) %>%
  separate(`Completion Rate`, into = c("numerator", "denominator"), sep = "/", convert = TRUE) %>%
  mutate(`Completion Rate %` = round(numerator / denominator,2) * 100,
          Average_Metres_Per_Run = round(as.numeric(`Run Metres`)/as.numeric(Runs),2)) %>%
  select(-numerator, -denominator) %>%
  mutate(across(c("Scores","Possession","Territory","Runs",
                  "Run Metres","Dummy Half Runs","Tackle Busts","Post Contact Metres","Offloads","Linebreaks",
                  "20m Restarts","Tackled In Opp 20","In Goal Escapes","Set Restarts Awarded","Tackles","Missed Tackles",
                  "Correct","Incorrect","Errors","Penalties Conceded","Kicks","Kick Metres",
                  "40/20s","20/40s","Attacking Kicks","Drop Outs","Forced Drop Outs","Kicks Dead",
                  "Completion Rate %",'Average_Metres_Per_Run'), ~ as.numeric(.))) %>%
  mutate(Season = case_when(
     startsWith(ID, "2024") ~ "2024",
     startsWith(ID, "2023") ~ "2023",
     startsWith(ID, "2022") ~ "2022",
     startsWith(ID, "2021") ~ "2021")) %>% 
  mutate('Game Type' = case_when(ID %in% c('2024-31-1','2024-30-1','2024-30-2',
                                           "2024-29-1","2024-29-2","2024-28-1",
                                           "2024-28-2","2024-28-3","2024-28-4",
                                           '2023-31-1','2023-30-2','2023-30-1',
                                           "2023-29-1","2023-29-2","2023-28-1",
                                           "2023-28-2","2023-28-3","2023-28-4",
                                           "2022-26-1","2022-26-2","2022-26-3",
                                           "2022-26-4","2022-27-1","2022-27-2",
                                           "2022-28-1","2022-28-2","2022-29-1",
                                           "2021-26-1","2021-26-2","2021-26-3",
                                           "2021-26-4","2021-27-1","2021-27-2",
                                           "2021-28-1","2021-28-2","2021-29-1") ~ 'Playoff',
                                 TRUE ~ 'Regular Season'))
```

```{r Full_Fast_NRLr, include=FALSE}
# Creating Full_Fast_NRLr

even_indices <- seq(2, nrow(NRL_seasons), by = 2)
odd_indices <- seq(1, nrow(NRL_seasons), by = 2)

# Subset the original dataframe using the even indices
NRL_away_teams <- NRL_seasons[even_indices, ]
NRL_home_teams <- NRL_seasons[odd_indices, ]

Fast_NRLr <- inner_join(NRL_home_teams,NRL_away_teams,by=c('ID','Season','Game Type'))
view(Fast_NRLr)

New_cols <- c("ID","Home Team","Home Score",
"Home Possession","Home Territory","Home Runs",
"Home Run Metres","Home Dummy Half Runs","Home Tackle Busts",
"Home Post Contact Metres", "Home Offloads", "Home Linebreaks",
"Home 20m Restarts","Home Tackled In Opp 20","Home In Goal Escapes",
"Home Awarded Set Restarts","Home Tackles","Home Missed Tackles",
"Home Correct Reviews","Home Incorrect Reviews","Home Errors",
"Home Penalties Conceded",  "Home Kicks","Home Kick Metres",
"Home 40/20s","Home 20/40s","Home Attacking Kicks",
"Home Drop Outs","Home Forced Drop Outs","Home Kicks Dead",
"Home Completion Rate","Home Average Metres Per Run",
"Season",'Game Type', "Away Team","Away Score",
"Away Possession","Away Territory","Away Runs",
"Away Run Metres","Away Dummy Half Runs","Away Tackle Busts",
"Away Post Contact Metres","Away Offloads","Away Linebreaks",
"Away 20m Restarts","Away Tackled In Opp 20","Away In Goal Escapes",
"Away Awarded Set Restarts", "Away Tackles","Away Missed Tackles",
"Away Correct Reviews","Away Incorrect Reviews","Away Errors",
"Away Penalties Conceded","Away Kicks","Away Kick Metres",
"Away 40/20s","Away 20/40s","Away Attacking Kicks",
"Away Drop Outs","Away Forced Drop Outs","Away Kicks Dead",
"Away Completion Rate","Away Average Metres Per Run") # "Game Type 1")

colnames(Fast_NRLr) <- New_cols
view(Fast_NRLr)

Fast_NRLr <- Fast_NRLr %>%
  relocate("ID","Home Team","Home Score","Away Team","Away Score",
"Home Possession","Home Territory",
"Away Possession","Away Territory",
"Home Runs","Home Run Metres","Away Runs",
"Away Run Metres","Home Dummy Half Runs","Away Dummy Half Runs",
"Home Tackle Busts","Away Tackle Busts",
"Home Post Contact Metres","Away Post Contact Metres","Home Average Metres Per Run",
"Away Average Metres Per Run","Home Offloads","Away Offloads",
"Home Linebreaks","Away Linebreaks",
"Home 20m Restarts","Away 20m Restarts","Home Completion Rate",
"Away Completion Rate","Home Tackled In Opp 20","Away Tackled In Opp 20",
"Home In Goal Escapes","Away In Goal Escapes","Home Awarded Set Restarts","Away Awarded Set Restarts",
"Home Tackles", "Away Tackles","Home Missed Tackles",
"Away Missed Tackles","Home Correct Reviews","Home Incorrect Reviews",
"Away Correct Reviews","Away Incorrect Reviews",
"Home Errors","Away Errors","Home Penalties Conceded","Away Penalties Conceded",
"Home Kicks","Away Kicks",
"Home Kick Metres","Away Kick Metres","Home 40/20s","Home 20/40s",
"Away 40/20s","Away 20/40s",
"Home Attacking Kicks","Away Attacking Kicks","Home Drop Outs",
"Away Drop Outs","Home Forced Drop Outs","Away Forced Drop Outs",
"Home Kicks Dead","Away Kicks Dead","Season") %>%
mutate(across(c("Home Score","Away Score","Home Possession","Home Territory",
                "Away Possession","Away Territory","Home Completion Rate",
                "Away Completion Rate","Home Runs","Home Run Metres","Away Runs",
                "Away Run Metres","Home Dummy Half Runs","Away Dummy Half Runs",
                "Home Tackle Busts","Away Tackle Busts","Home Post Contact Metres",
                "Away Post Contact Metres","Home Offloads","Away Offloads",
                "Home Linebreaks","Away Linebreaks",
                "Home 20m Restarts","Away 20m Restarts","Home Tackled In Opp 20",
                "Away Tackled In Opp 20",
                "Home In Goal Escapes","Away In Goal Escapes","Home Awarded Set Restarts",
                "Away Awarded Set Restarts",
                "Home Tackles", "Away Tackles","Home Missed Tackles",
                "Away Missed Tackles","Home Correct Reviews","Home Incorrect Reviews",
                "Away Correct Reviews","Away Incorrect Reviews",
                "Home Errors","Away Errors","Home Penalties Conceded","Away Penalties Conceded",
                "Home Kicks","Away Kicks",
                "Home Kick Metres","Away Kick Metres","Home 40/20s","Home 20/40s",
                "Away 40/20s","Away 20/40s",
                "Home Attacking Kicks","Away Attacking Kicks","Home Drop Outs",
                "Away Drop Outs","Home Forced Drop Outs","Away Forced Drop Outs",
                "Home Kicks Dead","Away Kicks Dead",
                "Home Average Metres Per Run","Away Average Metres Per Run"),
                 ~ as.numeric(.))) %>%
  mutate('Home result' = case_when(as.numeric(as.character(`Home Score`)) > `Away Score` ~ 'W',
                                   `Home Score` == `Away Score` ~ 'D',
                                   `Away Score` > `Home Score` ~ 'L'),
         'Away result' = case_when(`Home Score` > `Away Score` ~ 'L',
                                   `Home Score` == `Away Score` ~ 'D',
                                   `Away Score` > `Home Score` ~ 'W'),
         'Home Completion Difference' = `Home Completion Rate` - `Away Completion Rate`,
         'Away Completion Difference' = `Away Completion Rate` - `Home Completion Rate`,
         'Home Score Against' = `Away Score`,
         'Away Score Against' = `Home Score`) %>%
  mutate('Total runs' = `Home Runs` + `Away Runs`)
view(Fast_NRLr)

# Home and Away splits
Away_Fast_NRLr <- Fast_NRLr %>% select(ID,Season,`Away Team`,`Away Score`,`Away result`,
                                       `Away Possession`,`Away Territory`,`Away Runs`,
                                       `Away Run Metres`,`Away Dummy Half Runs`,`Away Tackle Busts`,
                                       `Away Post Contact Metres`,`Away Offloads`,`Away Linebreaks`,
                                       `Away 20m Restarts`,`Away Completion Rate`,`Away Tackled In Opp 20`,
                                       `Away In Goal Escapes`,`Away Awarded Set Restarts`,`Away Tackles`,
                                       `Away Missed Tackles`,`Away Correct Reviews`,`Away Incorrect Reviews`,
                                       `Away Errors`,`Away Penalties Conceded`,`Away Kicks`,
                                       `Away Kick Metres`,`Away 40/20s`,`Away 20/40s`,
                                       `Away Attacking Kicks`,`Away Drop Outs`,`Away Forced Drop Outs`,
                                       `Away Forced Drop Outs`,`Away Kicks Dead`,`Away Completion Difference`,
                                       `Away Score Against`,`Away Average Metres Per Run`,`Home Average Metres Per Run`,
                                       `Home Linebreaks`,`Total runs`) %>%
  mutate('Average run difference' = `Away Average Metres Per Run` - `Home Average Metres Per Run`,
         'Line break rate' = round(`Away Linebreaks`/`Away Runs`,2),
         'Line break differential' = `Away Linebreaks` - `Home Linebreaks`)

Home_Fast_NRLr <- Fast_NRLr %>% select(ID,Season,`Home Team`,`Home Score`,`Home result`,
                                       `Home Possession`,`Home Territory`,`Home Runs`,
                                       `Home Run Metres`,`Home Dummy Half Runs`,`Home Tackle Busts`,
                                       `Home Post Contact Metres`,`Home Offloads`,`Home Linebreaks`,
                                       `Home 20m Restarts`,`Home Completion Rate`,`Home Tackled In Opp 20`,
                                       `Home In Goal Escapes`,`Home Awarded Set Restarts`,`Home Tackles`,
                                       `Home Missed Tackles`,`Home Correct Reviews`,`Home Incorrect Reviews`,
                                       `Home Errors`,`Home Penalties Conceded`,`Home Kicks`,
                                       `Home Kick Metres`,`Home 40/20s`,`Home 20/40s`,
                                       `Home Attacking Kicks`,`Home Drop Outs`,`Home Forced Drop Outs`,
                                       `Home Forced Drop Outs`,`Home Kicks Dead`,`Home Completion Difference`,
                                       `Home Score Against`,`Home Average Metres Per Run`,
                                       `Away Average Metres Per Run`,`Away Linebreaks`,`Total runs`) %>%
  mutate('Average run difference' = `Home Average Metres Per Run` - `Away Average Metres Per Run`,
         'Line break rate' = round(`Home Linebreaks`/`Home Runs`,2),
         'Line break differential' = `Home Linebreaks` - `Away Linebreaks`)

Full_colnames <- c("ID","Season","Team","Scores","Result","Possession","Territory","Runs",
                   "Run Metres","Dummy Half Runs","Tackle Busts","Post Contact Metres",
                   "Offloads","Linebreaks","20m Restarts","Completion Rate",
                   "Tackled In Opp 20","In Goal Escapes",
                   "Awarded","Tackles","Missed Tackles","Correct","Incorrect","Errors","Penalties Conceded",
                   "Kicks","Kick Metres","40/20s","20/40s","Attacking Kicks","Drop Outs",
                   "Forced Drop Outs","Kicks Dead","Completion Difference","Score Against",
                   "Average Metres Per Run","Average Metres Per Run Conceded",
                   'Line breaks conceded','Total runs', 'Average run difference',
                   'Line break rate','Line break differential')

colnames(Home_Fast_NRLr) <- Full_colnames
colnames(Away_Fast_NRLr) <- Full_colnames

# Creating Full_Fast_NRLr ####
Full_Fast_NRLr <- rbind(Home_Fast_NRLr,Away_Fast_NRLr) %>%
  mutate(Team = str_to_title(Team))

Full_Fast_NRLr <- Full_Fast_NRLr %>%
  mutate('Game Type' = case_when(ID %in% c('2024-31-1','2024-30-2','2024-30-1',
                                           "2024-28-1","2024-28-2","2024-28-3",
                                           "2024-28-4","2024-29-1","2024-29-2",
                                           '2023-31-1','2023-30-2','2023-30-1',
                                           "2023-28-1","2023-28-2","2023-28-3",
                                           "2023-28-4","2023-29-1","2023-29-2",
                                           "2022-26-1","2022-26-2","2022-26-3",
                                           "2022-26-4","2022-27-1","2022-27-2",
                                           "2022-28-1","2022-28-2","2022-29-1",
                                           "2021-26-1","2021-26-2","2021-26-3",
                                           "2021-26-4","2021-27-1","2021-27-2",
                                           "2021-28-1","2021-28-2","2021-29-1") ~ "Playoff",
                                 TRUE ~ 'Regular Season'))
```

```{r Fast_NRLr_long, include=FALSE}
# Creating Long version of data for logistic regression

Fast_NRLr_long <- Fast_NRLr %>%
  pivot_longer(cols = c(`Home Score`, `Away Score`),
               names_to = "Location",
               values_to = "Score") %>%
  mutate(Result = ifelse(Location == "Home Score", `Home result`, `Away result`)) %>%
  select(Score, Result,`Home Team`,`Away Team`,ID,`Home result`,
         `Away result`) %>%
  mutate(Result = str_replace(Result,"W","1"),
         Result = str_replace(Result,"L","0")) %>%
  filter(Result %in% c("0","1")) %>%
  mutate(Result = as.numeric(as.character(Result)))
```

### Logistic regression

```{r logistic_regression,echo=FALSE,include=FALSE}
for_model <- glm(Fast_NRLr_long$Result~Fast_NRLr_long$Score,
             family = "binomial")
summary(for_model)
```

The plot below shows how scoring (the x-axis value) relates to the expected probability of a team winning (the y-axis value). If you hover over the blue line, they will see the score and corresponding probability of winning.

For example, if a team scores 21.5 points, that would give them almost a 50% chance of winning, whereas scoring a little over 11 points would give a team almost a 105 probability of winning, and scoring 31.8 points would give a team a probability of winning of a little over 90%.

```{r logistic_regression_plot,echo=FALSE}
Log_Regression_Plot <- Fast_NRLr_long %>%
  ggplot(aes(x=Score,y=Result)) +
  geom_point() +
  stat_smooth(method = "glm",se=FALSE,
              method.args = list(family='binomial')) +
  labs(title = 'Logistic regression - Scores vs Result')
ggplotly(Log_Regression_Plot)
```

## How do you win NRL games?

Outgaining your opponents in run metres significantly increases your chances of winning.

Over the 828 games, teams that had more run metres than their opponents won 652 times, or almost 79% of games.

```{r run_metre_difference,include=TRUE,echo=FALSE}
Full_Fast_NRLr %>%
  filter(`Run metre differential` > 0) %>%
  count(Result) %>%
  mutate(Percentage = round(n / sum(n) * 100,2)) %>%
  arrange(desc(Result)) %>%
  DT::datatable(caption = 'Results for Positive Run Metre Differential',
                options = list(dom = 't'))
```

The percentage at which you complete sets compared to an opponent, which we've called Completion Difference, significantly increases your chances of winning.

Over 794 games, teams that completed its sets at a higher rate than an opponent won almost 68% of games.

```{r completion_difference,include=TRUE,echo=FALSE}
# Completion difference 

Full_Fast_NRLr %>%
  filter(`Completion Difference` > 0) %>%
  count(Result) %>%
  mutate(Percentage = round(n / sum(n) * 100,2)) %>% 
  rename('Matches' = 'n') %>% 
  arrange(desc(Result)) %>% 
  DT::datatable(caption = 'Results for Completion Difference',
                options = list(dom = 't'))
```

Similarly, in the 754 matches where teams broke the opposition defensive line more than an opponent, the team with the positive line break differential won more than 75% of the time.

```{r LBdifferential,include=TRUE,echo=FALSE}
Full_Fast_NRLr %>%
  filter(`Line break differential` > 0) %>%
  count(Result) %>%
  mutate(Percentage = round(n / sum(n) * 100,2)) %>% 
  rename('Matches' = 'n') %>% 
  arrange(desc(Result)) %>% 
  DT::datatable(caption = 'Results for Line Break differential',
                options = list(dom = 't'))
```

In 439 games where a team had both a superior line break differential and Completion difference, they won almost 90% of the time.

```{r LBdifferential_CompletionDifference,include=TRUE,echo=FALSE}
Full_Fast_NRLr %>%
  filter(`Line break differential` > 0 & `Completion Difference` > 0) %>%
  count(Result) %>%
  mutate(Percentage = round(n / sum(n) * 100,2)) %>% 
  rename('Matches' = 'n') %>% 
  arrange(desc(Result)) %>% 
  DT::datatable(caption = 'Results for teams with superior Line Breaks & Completions',
                options = list(dom = 't'))
```

## Attack correlation/regression

```{r AttackCorrPlot,include=TRUE,echo=FALSE,fig.width=10}

Attack <- Full_Fast_NRLr %>%
   select(Scores,Possession,Territory,Runs,`Run Metres`,`Post Contact Metres`,
          `Linebreaks`,`Tackle Busts`,Offloads,Errors,`Tackled In Opp 20`,
          `Run Metres`,`Post Contact Metres`,`Completion Rate`,
          `Completion Difference`,`Average Metres Per Run`,
          `Average run difference`) %>%
   cor(use = "complete.obs") 
ggcorrplot(Attack,lab = TRUE,
           lab_size = 3,title = 'Attack Correlation Matrix')
```

## Defence Correlation Matrix

```{r DefenceCorrPlot,include=TRUE,echo=FALSE, fig.width=10}

Defence <- Full_Fast_NRLr %>%
  select(`Score Against`,Possession,Territory,Tackles,`Missed Tackles`,
         Errors,`Penalties Conceded`,`Completion Rate`,
         `Completion Difference`,`Average Metres Per Run Conceded`,
         `Line breaks conceded`) %>%
  cor(use = "complete.obs")
ggcorrplot(Defence,lab = TRUE,
           lab_size = 3,title = 'Defence Correlation Matrix')
```
